```{r}
# Load required packages
library(readr)
library(lubridate)
library(dplyr)

# Read the CSV file
df <- read.csv("D:/Dpa Project/DPA PROJECT/sentiment.csv")[, c(2, 8, 9)]

# Rename the columns if needed
colnames(df) <- c("Date", "Positive Sentiment", "Negative Sentiment")

# Convert the Date column to a date object
df$Date <- as_date(ymd_hms(df$Date))

# Set the Date column as the index and sort by index
df <- df %>% arrange(Date)

# Print the dimensions of the dataframe
print(dim(df))

# Print the first 5 rows
head(df, 5)
```

```{r}
str(df)
```

The first thing we need to resolve is how to score dates that have more than one Sentiment Score. This means there were several different articles parsed for that day. For instance, we can see below, September 8$^{th}$ had 4 different article parsed

```{r}
# Sort the dataframe by row names (assuming it's an index)
df_sorted <- df[order(rownames(df)), ]

# Subset the dataframe to rows 26 to 30 (inclusive)
df_subset <- df_sorted[26:30, ]
df_subset
```

There's a few ways we can handle this situation. We can take the Average of these values to give an overall sentiment for the day. A better way to average values to better represent orders of magnitude of each individual value is the Root Mean Square (RMS) which is simplyt $\sqrt{x_1^2 + x_2^2 + ... + x_n^2}$. And finally, we can simply sum the values. After performing this aggregations, let's then combine the negative and positive sentiment score to create some meaningful "Overall Sentiment" for each day. We will do this by simply summing the Positive Sentiment and Negative Sentiment

```{r}
head(df, 5)
```

```{r}
library(dplyr)

# group the data by Date
df <- df %>% 
  group_by(Date) %>% 

  # compute the mean, sum, and RMS of Positive Sentiment
  mutate(`Positive Sentiment Averaged` = mean(`Positive Sentiment`),
         `Positive Sentiment Summed` = sum(`Positive Sentiment`),
         `Positive Sentiment RMS` = sqrt(mean(`Positive Sentiment`^2))) %>% 

  # compute the mean, sum, and RMS of Negative Sentiment
  mutate(`Negative Sentiment Averaged` = mean(`Negative Sentiment`),
         `Negative Sentiment Summed` = sum(`Negative Sentiment`),
         `Negative Sentiment RMS` = sqrt(mean(`Negative Sentiment`^2))) %>% 
  
  # compute the overall sentiment
  mutate(`Overall Sentiment` = `Positive Sentiment` + `Negative Sentiment`) %>%
  
  # compute the mean, sum, and RMS of Overall Sentiment
  mutate(`Overall Sentiment Averaged` = mean(`Overall Sentiment`),
         `Overall Sentiment Summed` = sum(`Overall Sentiment`),
         `Overall Sentiment RMS` = sqrt(mean(`Overall Sentiment`^2))) %>% 
  
  # remove the grouping
  ungroup()

head(df)
```

```{r}
df <- df %>% 
  arrange(Date)

# select rows 25 to 26
df_subset <- df %>% 
  slice(25:35)

df_subset
```

I see the Mistake, there are multiple records for same date, these are duplicate records, we can summarise function to merge them into single one.

```{r}
library(dplyr)

# group the data by Date and summarize the columns
df <- df %>% 
  group_by(Date) %>% 

  summarize(`Positive Sentiment Averaged` = mean(`Positive Sentiment`),
            `Positive Sentiment Summed` = sum(`Positive Sentiment`),
            `Positive Sentiment RMS` = sqrt(mean(`Positive Sentiment`^2)),
            `Negative Sentiment Averaged` = mean(`Negative Sentiment`),
            `Negative Sentiment Summed` = sum(`Negative Sentiment`),
            `Negative Sentiment RMS` = sqrt(mean(`Negative Sentiment`^2)),
            `Overall Sentiment Averaged` = mean(`Positive Sentiment` + `Negative Sentiment`),
            `Overall Sentiment Summed` = sum(`Positive Sentiment` + `Negative Sentiment`),
            `Overall Sentiment RMS` = sqrt(mean((`Positive Sentiment` + `Negative Sentiment`)^2))) %>% 

  # remove the grouping
  ungroup()

# view the first few rows of the modified data frame
head(df)

```

```{r}
df <- df %>% 
  arrange(Date)

# select rows 25 to 26
df_subset <- df %>% 
  slice(26)

df_subset
```

```{r}
str(df)
```

### Historical Price Data

```{r}
library(data.table)


df2 <- read.csv("D:/Dpa Project/DPA PROJECT/historical_pricing.csv", header = TRUE)
names(df2)[1] <- "Date"
df2 <- df2[-1,]
df2$Date <- as.Date(df2$Date, format = '%m/%d/%y')
df2 <- df2[order(df2$Date),]
df2

```
```{r}
# Print the dimensions of the data frame
cat("shape of data frame: ", dim(df2), "\n")
cat("*************************************\n")
# Print the column names
cat("columns/variables/features in datasets:\n")
cat(names(df2), "\n")
cat("*************************************\n")

# Print the numerical columns
num_cols <- sapply(df2, is.numeric)
cat("Numerical columns in dataset:\n")
cat(names(df2[num_cols]), "\n")
cat("*************************************\n")

# Print the categorical columns
cat_cols <- sapply(df2, is.factor)
cat_cols <- cat_cols | sapply(df2, is.character)
cat("Categorical columns in dataset:\n")
cat(names(df2[cat_cols]), "\n")
```
The price data is actually numerical columns, We have to fix these, we fixed it using change format and rename column names.
```{r}
str(df2)
```
```{r}
colSums(is.na(df2))

# #no missing values present 
```
This will return a vector with the number of missing values in each column of the data frame. If the result shows all zeros, it means there are no missing values present in the data frame. If there are missing values, the corresponding columns will show non-zero values.


```{r}
# Count duplicate rows in the crypto_df data frame
sum(duplicated(df2))

#no duplicate values is present in dataset
```



```{r}

# Rename the columns
colnames(df2) <- c("Date", "XRP_Change","XRP_High","XRP_Low", "XRP_Open", "XRP_Price", "XRP_Vol", "BTC_Change", "BTC_High", "BTC_Low", "BTC_Open", "BTC_Price", "BTC_Vol", "ETH_Change", "ETH_High", "ETH_Low", "ETH_Open", "ETH_Price", "ETH_Vol")

head(df2)
```

```{r}
library(dplyr)
# Full Outer Join
df3 <- merge(df, df2, by.x = "Date",by.y="Date", all.x= TRUE,all.y=TRUE ,sort = TRUE)
df3 <- df3[-c(1:6),]
df3 <- df3[-1,]
df3[,-1] <- lapply(df3[,-1], function(x) ifelse(is.na(x), 0, x))
```

```{r}
df3$XRP_Change <- as.numeric(sub("%", "", df3$XRP_Change))
df3$XRP_High <- as.numeric(sub("%", "",df3$XRP_High))
df3$XRP_Low <- as.numeric(sub("%", "",df3$XRP_Low ))
df3$XRP_Open <- as.numeric(sub("%", "",df3$XRP_Open ))
df3$XRP_Price <- as.numeric(sub("%", "",df3$XRP_Price ))


df3$BTC_Change <- as.numeric(sub("%", "", df3$BTC_Change))
df3$BTC_High <- as.numeric(sub("%", "",df3$BTC_High))
df3$BTC_Low <- as.numeric(sub("%", "",df3$BTC_Low ))
df3$BTC_Open <- as.numeric(sub("%", "",df3$BTC_Open ))
df3$BTC_Price <- as.numeric(sub("%", "",df3$BTC_Price ))



df3$ETH_Change <- as.numeric(sub("%", "", df3$ETH_Change))
df3$ETH_High <- as.numeric(sub("%", "",df3$ETH_High))
df3$ETH_Low <- as.numeric(sub("%", "",df3$ETH_Low ))
df3$ETH_Open <- as.numeric(sub("%", "",df3$ETH_Open ))
df3$ETH_Price <- as.numeric(sub("%", "",df3$ETH_Price ))


df3$XRP_Vol <- as.numeric(sub("M", "", df3$XRP_Vol))/1000
df3$BTC_Vol <- as.numeric(sub("K", "", df3$BTC_Vol))
df3$ETH_Vol <- as.numeric(sub("K", "", df3$ETH_Vol))
```

```{r}
convert_thousands <- function(x) {
  x <- (x * 1000000) / 1000
  return(x)
}
```

```{r}
df3$XRP_Vol<- df3$XRP_Vol * 1000000 / 1000
df3$XRP_Change <- c(df3$XRP_Change[-1], 0)
df3$XRP_Price <- c(df3$XRP_Price[-1], 0)
```

```{r}
# view the fixed data frame
df3
```

```{r}
summary(df3)
```


```{r}
# generate the descriptive statistics with fivenum function
desc_stats <- apply(df3[,2:9], 2, fivenum)
desc_stats
# create a data frame from the statistics
desc_stats_df<- data.frame(t(desc_stats))
colnames(desc_stats_df) <- c("Min", "25%", "Median", "75%", "Max")
desc_stats_df$Mean <- apply(df3[,2:9], 2, mean)
desc_stats_df$SD <- apply(df3[,2:9], 2, sd)
desc_stats_df$Count <- apply(df3[,2:9], 2, function(x) sum(!is.na(x)))
desc_stats1<- data.frame(t(desc_stats_df))

# print the data frame
desc_stats1
```

```{r}
# function to compute statistics including missing values
my_stats <- function(x) {
  c(min(x, na.rm=TRUE), quantile(x, probs=c(0.25, 0.5, 0.75), na.rm=TRUE),
    max(x, na.rm=TRUE), sum(!is.na(x)))
}
# compute descriptive statistics
numeric_cols <- sapply(df3[, 11:(ncol(df3) - 1)], is.numeric)
desc_statss <- apply(df3[, 11:(ncol(df3) - 1)][, numeric_cols], 2, my_stats)

# create a data frame from the statistics
desc_stats_df1 <- data.frame(t(desc_statss))
colnames(desc_stats_df1) <- c("Min", "25%", "Median", "75%", "Max", "Count")
desc_stats_df1$Mean <- colMeans(df3[, 11:(ncol(df3) - 1)][, numeric_cols], na.rm=TRUE)
desc_stats_df1$SD <- apply(df3[, 11:(ncol(df3) - 1)][, numeric_cols], 2, sd, na.rm=TRUE)
desc_stats_df2 <- t(desc_stats_df1)

# print the data frame
library(tibble)
as_tibble(desc_stats_df2)
```

```{r}
# Remove rows with NA values
df_clean <- na.omit(desc_stats_df2)

# Create the correlation matrix
corr_matrix <- cor(df_clean)

# Select the columns
selected_columns <- corr_matrix[, c("XRP_Change", "XRP_Price")]

# Print the selected columns
print(selected_columns)

```

```{r}
library(ggplot2)
library(gridExtra)

# Assuming df3 is a data frame in R

# Get the first 9 column names
y_vars <- colnames(df3)[1:9]

# Create a list to store the individual plots
plot_list <- list()

# Loop through the y_vars and create a scatter plot for each
for (i in seq_along(y_vars)) {
  y_var <- y_vars[i]
  plot1 <- ggplot(df3, aes(x = XRP_Price, y = .data[[y_var]])) +
    geom_point() +
    theme_minimal() +
    xlab("XRP_Price") +
    ylab(y_var)
  plot2 <- ggplot(df3, aes(x = XRP_Change, y = .data[[y_var]])) +
    geom_point() +
    theme_minimal() +
    xlab("XRP_Change") +
    ylab(y_var)
  
  # Combine the plots for XRP_Price and XRP_Change for the current y_var
  combined_plot <- grid.arrange(plot1, plot2, ncol = 2)
  
  # Add the combined plot to the plot list
  plot_list[[i]] <- combined_plot
}

# Display the plots in a grid
do.call("grid.arrange", c(plot_list, ncol = 2))

```

```{r}
library(ggplot2)
library(gridExtra)

# Define the columns to use
x_vars <- c("XRP_Price", "XRP_Change")
y_vars <- c("BTC_High", "BTC_Low", "BTC_Price", "ETH_High", "ETH_Low", "ETH_Price")

# Create a list to store the individual plots
plot_list <- list()

# Loop through the y_vars and create a scatter plot for each
for (i in seq_along(y_vars)) {
  y_var <- y_vars[i]
  plot1 <- ggplot(df3, aes(x = .data[[x_vars[1]]], y = .data[[y_var]])) +
    geom_point() +
    theme_minimal() +
    xlab(x_vars[1]) +
    ylab(y_var)
  plot2 <- ggplot(df3, aes(x = .data[[x_vars[2]]], y = .data[[y_var]])) +
    geom_point() +
    theme_minimal() +
    xlab(x_vars[2]) +
    ylab(y_var)
  
  # Combine the plots for the x_vars for the current y_var
  combined_plot <- grid.arrange(plot1, plot2, ncol = 2)
  
  # Add the combined plot to the plot list
  plot_list[[i]] <- combined_plot
}

# Display the plots in a grid
do.call("grid.arrange", c(plot_list, ncol = 2))

```

```{r}
library(caret)
scaler <- preProcess(df3[,c('XRP_Price','BTC_Price','ETH_Price')], method = c("center", "scale"))
scaled_data <- predict(scaler, df3[,c('XRP_Price','BTC_Price','ETH_Price')])
scaled_data <- as.data.frame(scaled_data)
colnames(scaled_data) <- c('XRP','BTC','ETH')
rownames(scaled_data) <- rownames(df3)
head(scaled_data)
```

```{r}
library(ggplot2)

# Scale the data
scaled_data <- as.data.frame(scale(df3[, c("XRP_Price", "BTC_Price", "ETH_Price")]))
colnames(scaled_data) <- c("XRP", "BTC", "ETH")
scaled_data$Date <- df3$Date
scaled_data <- scaled_data[!is.na(scaled_data$Date), ]

# Plot the data
ax <- ggplot(scaled_data, aes(x = Date)) +
  geom_line(aes(y = XRP, colour = "XRP", linetype = "solid"), size = 1) +
  geom_line(aes(y = BTC, colour = "BTC", linetype = "solid"), size = 1) +
  geom_line(aes(y = ETH, colour = "ETH", linetype = "solid"), size = 1) +
  labs(x = "Date", y = "Normalized Price", title = "XRP - BTC - ETH PRICE") +
  scale_colour_manual(values = c("blue", "green", "red")) +
  theme_bw()

# Remove axes labels
ax <- ax + theme(axis.title.x = element_blank(), axis.title.y = element_blank())

# Remove ticks and tick labels
ax <- ax + theme(axis.text.x = element_blank(), axis.text.y = element_blank())
ax <- ax + theme(axis.ticks.x = element_blank(), axis.ticks.y = element_blank())

# Display the plot
print(ax)
which(is.na(scaled_data))
```

```{r}
library(caret)
scaler <- preProcess(df3[,c('BTC_High','BTC_Low','ETH_High','ETH_Low','XRP_Price')], method = "range")
data <- predict(scaler, df3[,c('BTC_High','BTC_Low','ETH_High','ETH_Low','XRP_Price')])
data <- as.data.frame(data, col.names = c('BTC_High','BTC_Low','ETH_High','ETH_Low','XRP_Price'), row.names = rownames(df3))
```

```{r}
library(ggplot2)

# plot BTC price with XRP price
ggplot(df3, aes(x = Date)) +
  geom_line(aes(y = BTC_High), color = 'green') +
  geom_line(aes(y = BTC_Low), color = 'red') +
  geom_line(aes(y = BTC_Price), color = 'black', linetype = 'dashed') +
  geom_ribbon(aes(ymin = BTC_Low, ymax = BTC_High), fill = 'red', alpha = 0.4) +
  labs(x = 'Date', y = 'Price', title = 'XRP Price vs BTC HLC') +
  theme_bw() +
  scale_y_continuous(sec.axis = sec_axis(~ ., name = 'XRP_Price', breaks = seq(0, 4, 0.5))) +
  theme(legend.position = 'bottom')

# plot ETH price with XRP price
ggplot(df3, aes(x = Date)) +
  geom_line(aes(y = ETH_High), color = 'green') +
  geom_line(aes(y = ETH_Low), color = 'red') +
  geom_line(aes(y = ETH_Price), color = 'black', linetype = 'dashed') +
  geom_ribbon(aes(ymin = ETH_Low, ymax = ETH_High), fill = 'red', alpha = 0.4) +
  labs(x = 'Date', y = 'Price', title = 'XRP Price vs ETH HLC') +
  theme_bw() +
  scale_y_continuous(sec.axis = sec_axis(~ ., name = 'XRP_Price', breaks = seq(0, 4, 0.5))) +
  theme(legend.position = 'bottom')

```

```{r}
lm_p <- lm(XRP_Price ~ `Positive Sentiment Summed`, data = df3)
lm_n <- lm(XRP_Price ~ `Negative Sentiment Summed`, data = df3)

y_p <- predict(lm_p)
y_n <- predict(lm_n)
```

```{r}
library(ggplot2)


ggplot(df3, aes(x = `Negative Sentiment RMS`, y = XRP_Price)) +
  geom_point(aes(color = "Negative Sentiment"), alpha = 0.4, shape = 20) +
  geom_point(aes(x = `Positive Sentiment RMS`, y = XRP_Price, color = "Positive Sentiment"), alpha = 0.4, shape = 20) +
  scale_color_manual(values = c("Negative Sentiment" = "red", "Positive Sentiment" = "green")) +
  labs(x = "Negaitve - Positve Sentiment RMS", y = "XRP Price", title = "Sentiment Analysis vs. XRP Price") +
  stat_ellipse(aes(x0 = -0.3, y0 = 1.5, a = 0.2, b = 2.8, angle = 348), type = "norm", level = 0.95, geom = "polygon", color = "gray", fill = "gray", alpha = 0.2, linetype = 0, size = 2) +
  stat_ellipse(aes(x0 = 0.3, y0 = 1.9, a = 0.2, b = 2.7, angle = 349), type = "norm", level = 0.95, geom = "polygon", color = "gray", fill = "gray", alpha = 0.2, linetype = 0, size = 2) +
  theme_minimal()

ggplot(df3, aes(x = `Negative Sentiment Summed`, y = XRP_Price)) +
  geom_point(aes(color = "Negative Sentiment"), alpha = 0.4, shape = 20) +
  geom_point(aes(x = `Positive Sentiment Summed`, y = XRP_Price, color = "Positive Sentiment"), alpha = 0.4, shape = 20) +
  scale_color_manual(values = c("Negative Sentiment" = "red", "Positive Sentiment" = "green")) +
  labs(x = "Negative - Positive Sentiment Summed", y = "XRP Price", title = "Sentiment Analysis vs. XRP Price") +
  stat_smooth(aes(x = `Positive Sentiment Summed`, y = XRP_Price), method = "lm", color = "green") +
  stat_smooth(aes(x = `Negative Sentiment Summed`, y = XRP_Price), method = "lm", color = "red") +
  theme_minimal()


```

```{r}

library(ggplot2)
library(gridExtra)

# create the first plot
p1 <- ggplot(df3, aes(x = Date)) + 
  geom_line(aes(y = XRP_Price, colour = "Price"), size = 0.5) +
  geom_ribbon(aes(ymin = 0, ymax = XRP_Price, fill = "Price"), alpha = 0.4) +
  geom_line(aes(y = `Positive Sentiment RMS`, colour = "Positive Sentiment RMS"), size = 0.5) +
  geom_ribbon(aes(ymin = 0, ymax = `Positive Sentiment RMS`, fill = "Positive Sentiment RMS"), alpha = 0.4) +
  geom_line(aes(y = `Negative Sentiment RMS`, colour = "Negative Sentiment RMS"), size = 0.5) +
  geom_ribbon(aes(ymin = 0, ymax = `Negative Sentiment RMS`, fill = "Negative Sentiment RMS"), alpha = 0.4) +
  scale_fill_manual(values = c("Price" = "blue", "Positive Sentiment RMS" = "green", "Negative Sentiment RMS" = "red")) +
  scale_color_manual(values = c("Price" = "black", "Positive Sentiment RMS" = "black", "Negative Sentiment RMS" = "black")) +
  labs(x = "", y = "Price / Sentiment RMS", color = "", fill = "") +
  ggtitle("Price with Sentiment RMS") +
  theme_classic() +
  theme(plot.title = element_text(size = 18, face = "bold"),
        legend.position = "bottom",
        legend.key.size = unit(0.8, "cm"),
        legend.text = element_text(size = 14),
        legend.title = element_blank(),
        axis.text = element_text(size = 14),
        axis.title = element_text(size = 16))

# create the second plot
p2 <- ggplot(df3, aes(x = Date)) + 
  geom_line(aes(y = XRP_Price, colour = "Price"), size = 0.5) +
  geom_ribbon(aes(ymin = 0, ymax = XRP_Price, fill = "Price"), alpha = 0.4) +
  geom_line(aes(y = `Positive Sentiment Summed`, colour = "Positive Sentiment Summed"), size = 0.5) +
  geom_ribbon(aes(ymin = 0, ymax = `Positive Sentiment Summed`, fill = "Positive Sentiment Summed"), alpha = 0.4) +
  geom_line(aes(y = `Negative Sentiment Summed`, colour = "Negative Sentiment Summed"), size = 0.5) +
  geom_ribbon(aes(ymin = 0, ymax = `Negative Sentiment Summed`, fill = "Negative Sentiment Summed"), alpha = 0.4) +
  scale_fill_manual(values = c("Price" = "blue", "Positive Sentiment Summed" = "green", "Negative Sentiment Summed" = "red")) +
  scale_color_manual(values = c("Price" = "black", "Positive Sentiment Summed" = "black", "Negative Sentiment Summed" = "black")) +
  labs(x = "", y = "Price / Sentiment Summed", color = "", fill = "") +
  ggtitle("Price with Sentiment Summed") +
  theme_classic() +
  theme(plot.title = element_text(size = 18, face = "bold"),
        legend.position = "bottom",
        legend.key.size = unit(0.8, "cm"),
        legend.text = element_text(size = 14),
        legend.title = element_blank(),
        axis.text = element_text(size = 14),
        axis.title = element_text(size = 16))
grid.arrange(p1, ncol = 1, heights = c(2, 1))
 grid.arrange(p2, ncol = 1, heights = c(2, 1))
 
```

Again, we can see that Sentiment RMS doesn't translate very well to a change in XRP price like Sentiment Summed does. At XRP's highest Price point, Sentiment RMS has a value no higher than when XRP's price point was close to it's lowest. Sentiment Summed however, does reflect a huge leap in XRP Price at the same instance Sentiment explodes as well. However take note that these sentiments are purely day to day sentiments. But when an investor is investing, they wouldn't only look at articles on the day they were investing. They would research days, and sometimes weeks prior to determine sentiment and performance. Therefore, a days sentiment should also reflect previous days' sentiment as well. This will not only help smooth out the sentiment curve, but take into account for all 0 values of sentiment that the sentiment might not actually be 0. A value of 0 can simply mean that no article was written on that day, so Sentiment from a previous day would simply roll into that day. If There had been extremely high sentiment for the past week, but today there were no articles written on XRP, that doesn't mean that Sentiment on the coin has dropped within the market, it just simply means that no article was written today on the coin. We must account for this. It's not unreasonable to speculate that an investor might look at a coins weekly performance when determining strategy. Therefore, let's perform a 5 day rolling average for sentiment. When averaging the values however, yesterday's value being averaged into today will be it's original value, not it's rolling average value from when the 5 day average was performed on it. For example if we were performing a 3 day rolling average:

Day 20 - Original Value: 2 \| 3-Day Average: 2.7

Day 21 - Original Value: 1 \| 3-Day Average: 2.5

Day 23 - Original Value: 4 \| 3-Day Average: 3.5

Notice from day 3, it got it's 3 day average from the original values of the previous 2 days, not their 3 Day Average

```{r}
boxfilter <- function(df, col) {
  # create copies so as to not overwrite original dataframe
  c <- df[[col]]
  d <- c
  # overwrite values in d with rolling averages of the last 5 values in c
  for (i in 5:length(c)) {
    d[i] <- mean(c((c[(i-4):i])))
  }
  # create new column to write to dataframe
  column <- paste0(strsplit(col, ' ')[[1]][length(strsplit(col, ' ')[[1]])], '_box_filter')
  # set column equal to averaged sentiments within the dataframe
  df[column] <- d
  # return original dataframe with new column
  return(df)
}


```

```{r}
df4 <- df3
df4 = boxfilter(df4, 'Overall Sentiment RMS')
df4 = boxfilter(df4, 'Overall Sentiment Summed')
df4 = boxfilter(df4, 'Overall Sentiment Averaged')
```

```{r}
library(ggplot2)
library(gridExtra)

# create the first plot
p1 <- ggplot(df3, aes(x = Date)) + 
  geom_line(aes(y = XRP_Price, colour = "Price"), size = 0.5) +
  geom_ribbon(aes(ymin = 0, ymax = XRP_Price, fill = "Price"), alpha = 0.4) +
  geom_line(aes(y = `Positive Sentiment RMS`, colour = "Positive Sentiment RMS"), size = 0.5) +
  geom_ribbon(aes(ymin = 0, ymax = `Positive Sentiment RMS`, fill = "Positive Sentiment RMS"), alpha = 0.4) +
  geom_line(aes(y = `Negative Sentiment RMS`, colour = "Negative Sentiment RMS"), size = 0.5) +
  geom_ribbon(aes(ymin = 0, ymax = `Negative Sentiment RMS`, fill = "Negative Sentiment RMS"), alpha = 0.4) +
  scale_fill_manual(values = c("Price" = "blue", "Positive Sentiment RMS" = "green", "Negative Sentiment RMS" = "red")) +
  scale_color_manual(values = c("Price" = "black", "Positive Sentiment RMS" = "black", "Negative Sentiment RMS" = "black")) +
  labs(x = "", y = "Price / Sentiment RMS", color = "", fill = "") +
  ggtitle("Price with Sentiment RMS") +
  theme_classic() +
  theme(plot.title = element_text(size = 18, face = "bold"),
        legend.position = "bottom",
        legend.key.size = unit(0.8, "cm"),
        legend.text = element_text(size = 14),
        legend.title = element_blank(),
        axis.text = element_text(size = 14),
        axis.title = element_text(size = 16))
grid.arrange(p1, ncol = 1, heights = c(2, 1))

# create the second plot
par(mfrow = c(2,1), mar = c(4, 4, 2, 1), oma = c(0, 0, 2, 0))
plot(df4$XRP_Price, type = "l", col = "black", lwd = 0.5, xaxt = "n", xlab = "", ylab = "Price / RMS Sentiment")
polygon(c(1:length(df4$XRP_Price), length(df4$XRP_Price):1), c(rep(0, length(df4$XRP_Price)), rev(df4$XRP_Price)), col = "blue", border = NA, alpha = 0.4)
lines(df4$Summed_box_filter, col = "black", lwd = 0.5)
polygon(c(1:length(df4$Summed_box_filter), length(df4$Summed_box_filter):1), c(rep(0, length(df4$Summed_box_filter)), rev(df4$Summed_box_filter)), col = ifelse(df4$Summed_box_filter >= 0, "green", "red"), border = NA, alpha = 0.4)
title("Filtered RMS Sentiment with Price")
mtext("Price / RMS Sentiment", side = 2, line = 2, cex = 0.8)
axis(1, at = seq(1, nrow(df4), by = 30), labels = format(as.Date(df4$index)[seq(1, nrow(df4), by = 30)], "%b %d"))

```

```{r}
# Check for non-numeric values and convert to numeric
df4[] <- lapply(df4, as.numeric)

# Remove rows with missing data
df4_complete <- df4[complete.cases(df4),]

# Convert df4_complete to a numeric matrix
df4_numeric <- as.matrix(df4_complete)

# Compute the correlation matrix
corr <- cor(df4_numeric)

# Extract the correlation coefficient between "XRP_Price" and "Summed_box_filter"
corr_coef <- corr[["XRP_Price", "Summed_box_filter"]]
corr_coef 
```

```{r}
# Assuming 'df4' is already available as a data frame in R

# Load required packages
library(ggplot2)
library(dplyr)

# Create a new column 'colors' based on the 'XRP_Vol' column values
df4 <- df4 %>%
  mutate(colors = ifelse(lag(XRP_Vol, default = first(XRP_Vol)) < XRP_Vol, "g", "r"))

# Create the main plot
main_plot <- ggplot(df4, aes(x = Date)) +
  geom_line(aes(y = XRP_Price), color = "black", size = 0.5) +
  geom_area(aes(y = XRP_Price), fill = "blue", alpha = 0.4) +
  geom_line(aes(y = Summed_box_filter), color = "black", size = 0.5) +
  geom_area(aes(y = ifelse(Summed_box_filter >= 0, Summed_box_filter, 0)), fill = "green") +
  geom_area(aes(y = ifelse(Summed_box_filter <= 0, Summed_box_filter, 0)), fill = "red") +
  theme_void() +
  theme(plot.margin = margin(0, 0, 0, 0))

# Create the volume bar plot
volume_plot <- ggplot(df4, aes(x = Date, y = XRP_Vol, fill = colors)) +
  geom_col() +
  scale_fill_manual(values = c("g" = "green", "r" = "red")) +
  theme_void() +
  theme(plot.margin = margin(0, 0, 0, 0))

# Combine the main plot and volume bar plot using cowplot
library(cowplot)
combined_plot <- plot_grid(main_plot, volume_plot, nrow = 2, rel_heights = c(3, 1))

# Display the combined plot
print(combined_plot)



```

```{r}
rolling_average <- function(df) {
  a <- df
  c <- a$XRP_Price
  five <- a$XRP_Change
  five[1] <- 0
  ten <- five
  fifteen <- five
  twenty <- five
  thirty <- five
  
  for (i in 2:length(five)) {
    if (i < 5) {
      five[i] <- (c[i] - c[1]) / c[1]
    } else {
      five[i] <- (c[i] - c[i-4]) / c[i-4]
    }
  }
  
  for (i in 2:length(ten)) {
    if (i < 10) {
      ten[i] <- (c[i] - c[1]) / c[1]
    } else {
      ten[i] <- (c[i] - c[i-9]) / c[i-9]
    }
  }
  
  for (i in 2:length(fifteen)) {
    if (i < 15) {
      fifteen[i] <- (c[i] - c[1]) / c[1]
    } else {
      fifteen[i] <- (c[i] - c[i-14]) / c[i-14]
    }
  }
  
  for (i in 2:length(twenty)) {
    if (i < 20) {
      twenty[i] <- (c[i] - c[1]) / c[1]
    } else {
      twenty[i] <- (c[i] - c[i-19]) / c[i-19]
    }
  }
  
  for (i in 2:length(thirty)) {
    if (i < 30) {
      thirty[i] <- (c[i] - c[1]) / c[1]
    } else {
      thirty[i] <- (c[i] - c[i-29]) / c[i-29]
    }
  }
  
  data <- data.frame('five day percent change' = five, 'ten day percent change' = ten , 'fifteen day percent change' = fifteen, 'twenty day percent change' = twenty, 'thirty day percent change' = thirty, 'price' = a$XRP_Price)
  
  return(data)
}



```

```{r}
data <- rolling_average(df4)
head(data, 30)
```

```{r}
library(ggplot2)
# Assuming your data frame is named "data"

# Set the index as a separate column
data$index <- seq_along(data$price)

# Create a function to generate the plots
generate_plot <- function(col_name) {
  ggplot(data, aes(x = index)) +
    geom_line(aes(y = price), color = "black") +
    geom_area(aes(y = price), fill = "blue", alpha = 0.4) +
    geom_line(aes(y = data[[col_name]]), color = "black") +
    geom_area(aes(y = pmax(data[[col_name]], 0)), fill = "green", alpha = 0.4) +
    geom_area(aes(y = pmin(data[[col_name]], 0)), fill = "red", alpha = 0.4) +
    labs(title = col_name, y = "Price / Price Percent Change") +
    theme_minimal()
}

# Generate the plots
idx <- c("five.day.percent.change", "ten.day.percent.change", "fifteen.day.percent.change",
         "twenty.day.percent.change", "thirty.day.percent.change")

for (i in idx) {
  print(generate_plot(i))
}


```

```{r}

library(ggplot2)
library(ggcorrplot)

# Remove rows and columns with missing values
data_complete <- na.omit(data)

# Compute the correlation matrix
cor_matrix <- cor(data_complete)

# Create the correlation heatmap with correlation coefficients
ggcorrplot(cor_matrix, hc.order = TRUE, type = "lower", 
           lab = TRUE, lab_size = 3,
           colors = c("green", "white", "pink"),
           title = "Correlation Heatmap",
           ggtheme = ggplot2::theme_minimal())

```

```{r}

library(tidyr)

library(dplyr)

# Assuming 'data' and 'df3' are already available as data frames in R

a <- data %>% 
  mutate(row_index = row_number()) %>% 
  select(row_index, "ten.day.percent.change", "thirty.day.percent.change")

b <- df3 %>% 
  mutate(row_index = row_number()) %>% 
  full_join(a, by = "row_index")

# Select only numeric columns
b_numeric <- b %>% select_if(is.numeric)

corr <- cor(b_numeric, use = "complete.obs", method = "pearson")
filtered_corr <- corr[corr["ten.day.percent.change", ] > 0.3 | corr["thirty.day.percent.change", ] > 0.3, ]
filtered_corr <- filtered_corr %>% as_tibble(rownames = "variable") %>%
  filter(variable == "ten.day.percent.change" | variable == "thirty.day.percent.change") %>%
  drop_na()

filtered_corr

```

```{r}
library(dplyr)

filtered_corr <- corr[corr["XRP_Price", ] > 0.3, c("XRP_Price", "thirty.day.percent.change")]

filtered_corr

```

```{r}
library(dplyr)

# Assuming 'b' is already available as a data frame in R

bins_result <- cut(b$XRP_Price, breaks = 50, labels = FALSE, include.lowest = TRUE)
bins <- attr(bins_result, "breaks")

# Print bins
cat("Bins:", "\n")
print(bins)

# Add the bins_result as a new column to the data frame
b_with_bins <- b %>% mutate(XRP_Price_Bins = bins_result)

# Print the first 30 elements of the binned data
cat("Binned Data:", "\n")
print(head(b_with_bins[, c("Date", "XRP_Price_Bins")], 30))

```

```{r}
# Load the required libraries
library(ggplot2)
library(dplyr)

# Assuming 'b_with_bins' is already available as a data frame in R

# Create the ggplot object with the binned price (XRP_Price_Bins) and the Date
p <- ggplot(b_with_bins, aes(x = Date)) +
  geom_line(aes(y = XRP_Price_Bins), color = "blue", size = 1.5) +
  labs(y = "Bin Number", title = "Binned Price Ranges") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

# Add the XRP_Price on a secondary axis
p <- p + geom_line(aes(y = XRP_Price * 10), color = "red", linetype = "dashed", size = 1) +
  scale_y_continuous(sec.axis = sec_axis(~./10, name = "XRP Price"))

# Display the plot
print(p)

```

```{r}
# Assuming 'b_with_bins' is already available as a data frame in R

# Create a new column 'binned_price' with the integer values of 'bins_result'
b_with_bins$binned_price <- as.integer(bins_result)

# Select only numeric columns
numeric_columns <- sapply(b_with_bins, is.numeric)
b_with_bins_numeric <- b_with_bins[, numeric_columns]

# Calculate the correlation
corr_matrix <- cor(b_with_bins_numeric, use = "complete.obs", method = "pearson")

# Filter the correlation matrix for values greater than 0.3
filtered_corr <- corr_matrix
filtered_corr[filtered_corr <= 0.3] <- NA

# Subset the correlation matrix to only include 'binned_price' and 'XRP_Price' columns
filtered_corr <- subset(filtered_corr, select = c("binned_price", "XRP_Price"))

# Drop rows with all NA values
filtered_corr <- filtered_corr[!apply(is.na(filtered_corr), 1, all),]

# Display the filtered correlation matrix
print(filtered_corr)


```

```{r}
# Assuming 'b' is already available as a data frame in R

# Save the DataFrame 'b' to a CSV file
write.csv(b, file = "XRP_Data.csv", row.names = FALSE)

```





